---
title: "The BCB420.2019.STRINGnetwork: contruct correlation network for a system"
author:
- name: Han Zhang
  email: helennn.zhang@utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  This Vignette shows the package STRINGnetwork to obtain gene networks from the system PHYLA
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{BCB420.2019.ESA Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
- id: steipe2019
  title: 'rptPlus: R Package Template - enhanced'
  author:
  - family: Steipe
    given: Boris
    orcid: 0000-0002-1134-6758
  container-title: '[R package](https://github.com/hyginn/BCB420.2019.ESA)'
  URL: 'https://github.com/hyginn/BCB420.2019.ESA'
---


<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# knitr knits in a new session with an empty global workspace after setting its
# working directory to ./vignettes. To make your package functions available in
# the vignette, you have to load the library. The following two lines should
# accomplish this without manual intervention:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)

```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
There are many links and references in this document. If you find anything here
ambiguous, inaccurate, outdated, incomplete, or broken, please
[file an issue](https://github.com/hyginn/BCB420.2019.ESA/issues)!
</div>

&nbsp;

# About this Vignette

This sample Vignette explains the BCB420.2019.STRINGnetwork package and guide through how to use this tool.
This package provides the visualization of gene correlation within a system.

The core purposes of the tool:
1. To help study a system.
2. To provide an overview correlation.
3. Return network graphs on selected genes.

# Example
```
produceGeneNetwork = function(sys, gene="ATP2A3") {
  StringURL <- paste0("http://steipe.biochemistry.utoronto.ca/abc/assets/",
                      "STRINGedges-2019-03-14.RData")
  load(url(StringURL)) # loads STRING edges object
  
  # fetch data
  myDB <- fetchData("SysDB")
  geneSet <- SyDBgetSysSymbols(fetchData("SysDB"), sys)
  
  # 1.query a gene network
  # gene = "ATP2A3" #randomly select a gene
  # find the genes that is attached to the ATP2A3 by edges
  geneNetwork = STRINGedges[STRINGedges$a == gene|STRINGedges$b==gene ,]
  
  # get the number of genes in the list
  rowcount = dim(geneNetwork)[1]
  networkdata = c()
  # append all the genes to the list
  for(i in 1:rowcount){
    networkdata = append(networkdata,geneNetwork[i,]$a)
    networkdata = append(networkdata,geneNetwork[i,]$b)
  }
  # add edges between genes
  g = igraph::graph(edges=networkdata, directed=F)
  # plot the graph
  plot(g,edge.arrow.size=.2, vertex.color="red", vertex.size=2,
       vertex.frame.color="gray", vertex.label.color="black",
       vertex.label.cex=0.6, vertex.label.dist=1, edge.curved=0.1)
}
```
# Further reading
- https://github.com/hyginn/BCB420.2019.STRING
- https://github.com/hannzhang/BCB420.2019.ESA/blob/master/R/network.R


# Session Info

This release of the `BCB420.2019.STRINGnetwork` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

R version 3.5.1 (2018-07-02)
Platform: x86_64-apple-darwin15.6.0 (64-bit)
Running under: macOS  10.14.3

Matrix products: default
BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/3.5/Resources/lib/libRlapack.dylib

locale:
[1] en_CA.UTF-8/en_CA.UTF-8/en_CA.UTF-8/C/en_CA.UTF-8/en_CA.UTF-8

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] rmarkdown_1.10 knitr_1.20     testthat_2.0.1

loaded via a namespace (and not attached):
 [1] Rcpp_1.0.0         digest_0.6.18      rprojroot_1.3-2    R6_2.3.0          
 [5] backports_1.1.2    magrittr_1.5       evaluate_0.12      rlang_0.3.0.1     
 [9] tools_3.5.1        igraph_1.2.2       yaml_2.2.0         compiler_3.5.1    
[13] pkgconfig_2.0.2    BiocManager_1.30.4 htmltools_0.3.6 

<!-- the bibliography appears at the end of the page -->
# References
- https://github.com/hyginn/BCB420.2019.STRING

<!-- End -->
