---
title: "BCB420.2019.ESA-Vignette-Yuhan.Rmd"
author: "Yuhan Zhang"
date: "2019/3/25"
- name: Yuhan Zhang
  affiliation: 
  - https://orcid.org/0000-0003-1275-7951
  - University of Toronto
  email: yuhan.zhang@utoronto.ca
date: "`r Sys.Date()`"
abstract: >
This function will take a set of genes and collect their expression correlation and 
GO semantic similarity for each pair of genes
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{BCB420.2019.ESA Vignette Yuhan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
- id: steipe2019
  title: 'rptPlus: R Package Template - enhanced'
  author:
  - family: Steipe
    given: Boris
    orcid: 0000-0002-1134-6758
  container-title: '[R package](https://github.com/hyginn/BCB420.2019.ESA)'
  URL: 'https://github.com/hyginn/BCB420.2019.ESA'
---

<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# knitr knits in a new session with an empty global workspace after setting its
# working directory to ./vignettes. To make your package functions available in
# the vignette, you have to load the library. The following two lines should
# accomplish this without manual intervention:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)
```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
There are many links and references in this document. If you find anything here
ambiguous, inaccurate, outdated, incomplete, or broken, please
[file an issue](https://github.com/hyginn/BCB420.2019.ESA/issues)!
</div>

&nbsp;

# About this Vignette

This function takes a set of genes with hgnc symbol, calculating their correlations from expression files using corGenes function provided from Dr Boris. Calculate Go semantic similarity for gene products by GoSem function. Plot graph and build linear model of these two variables to show the relationship between gene correlation and GO similarity.

# Session Info

This release of the `BCB420.2019.ESA` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

<!-- the bibliography appears at the end of the page -->
# References


<!-- End -->
