---
title: "The BCB420.2019.ESA Vignette"
author:
- name: Boris Steipe
  affiliation: 
  - http://orcid.org/0000-0002-1134-6758
  - University of Toronto
  email: boris.steipe@utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  This Vignette is a placeholder for global package documentation.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{BCB420.2019.ESA Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
- id: steipe2019
  title: 'rptPlus: R Package Template - enhanced'
  author:
  - family: Steipe
    given: Boris
    orcid: 0000-0002-1134-6758
  container-title: '[R package](https://github.com/hyginn/BCB420.2019.ESA)'
  URL: 'https://github.com/hyginn/BCB420.2019.ESA'
---

<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# knitr knits in a new session with an empty global workspace after setting its
# working directory to ./vignettes. To make your package functions available in
# the vignette, you have to load the library. The following two lines should
# accomplish this without manual intervention:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)
```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
There are many links and references in this document. If you find anything here
ambiguous, inaccurate, outdated, incomplete, or broken, please
[file an issue](https://github.com/hyginn/BCB420.2019.ESA/issues)!
</div>

&nbsp;

# About this Vignette

This sample Vignette explains the BCB420.2019.ESA package. See the `rptPlus` package vignette for details of using .Rmd for vignettes and vignette formatting.


# Further reading
<a name="a1"></a>

- TBD

# Session Info

This release of the `BCB420.2019.ESA` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

<!-- the bibliography appears at the end of the page -->
# References




<!-- End -->
