---
title: "STRINGExplore: Interactive Gene and Protein Action Network"
author:
- name: Gabriela Morgenshtern
  affiliation: 
  - http://orcid.org/0000-0003-4762-8797
  - University of Toronto
  email: g.morgenshtern@mail.utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  This Vignette demonstrates how to use the STRINGExplore package to obtain an interactive network from high confidence STRING scores of a curated gene system.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{BCB420.2019.ESA Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
- id: steipe2019
  title: 'rptPlus: R Package Template - enhanced'
  author:
  - family: Steipe
    given: Boris
    orcid: 0000-0002-1134-6758
  container-title: '[R package](https://github.com/hyginn/BCB420.2019.ESA)'
  URL: 'https://github.com/hyginn/BCB420.2019.ESA'
---

<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# knitr knits in a new session with an empty global workspace after setting its
# working directory to ./vignettes. To make your package functions available in
# the vignette, you have to load the library. The following two lines should
# accomplish this without manual intervention:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)

```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
There are many links and references in this document. If you find anything here
ambiguous, inaccurate, outdated, incomplete, or broken, please
[file an issue](https://github.com/hyginn/BCB420.2019.ESA/issues)!
</div>

&nbsp;

# About this Vignette

This sample Vignette explains the BCB420.2019.STRINGExplore package. It aims to provide an instructive tutorial on the practical uses of the STRINGExplore tool and discuss applications and the interpretation of results.

The purpose of STRINGExplore is in the name-- it is exploratory. It aims to coordinate information available on disparate parts of the STRING database in a way that is visually pleasing and descriptive. These aims are achieved through the generation of a high-confidence gene interaction network and the leveraging of this visualization's tooltip system.

At its core, the aims of the descriptiveness of the system are as follows:
1. To inspire interaction
2. To develop the end-user's intuition for future hypotheses 
3. To be easy to understand and navigate
4. To ameliorate the current state of publication images (FIND EX)

## What's wrong with igraph?

STRINGExplore was created out of frustration about the complicated nature of setting up an igraph network in a visually pleasing and easy-to-analyze way. In regards to explorations of gene-gene interaction, igraph requires the user to be aware of the nature of their data before plotting it. With STRINGExplore, the nature of discovering the existence and strength of gene-gene interactions 

# Example usage
- add functionality for grouping by type of interaction: currently limited by the capabilities of grouping nodes in visNetwork ( Custom option. Character or a named list. Add a multiple selection based on column of node data.frame creating an HTML select element. Not available for DOT and Gephi.) Would require multiples of a single node to exist for you to be able to select by a "Protein Action". Thus the tooltip with sentence-based protein action information
- ref to above: https://www.rdocumentation.org/packages/visNetwork/versions/2.0.5/topics/visOptions
=======

# Saving and publishing your network
- add functionality for STRINGExplore.save() using https://rdrr.io/cran/visNetwork/man/visSave.html
- this saves the HTML of the network and all the dependencies of this analysis. Use this HTML on your research site for some snazzy interactive visualization axn

# On input data and preparation
# Further reading (maybe move to references?)
<a name="a1"></a>
- http://kbroman.org/pkg_primer/pages/vignettes.html
- https://rstudio-pubs-static.s3.amazonaws.com/157501_93a72a58ec614946901e10edf78c1384.html
- https://github.com/gabmorg/bch411-2018/blob/master/2018-10-17_Graphs-and-Networks_v01.R
- https://datastorm-open.github.io/visNetwork/options.html
- http://kateto.net/wp-content/uploads/2017/06/Polnet%202017%20R%20Network%20Visualization%20Workshop.pdf
- https://github.com/hyginn/BCB420.2019.STRING
- https://datastorm-open.github.io/visNetwork/options.html

# Session Info

This release of the `BCB420.2019.STRINGExplore` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

<!-- the bibliography appears at the end of the page -->
# References


<!-- End -->
