---
title: "Data Details - TBD"
author:
- name: Boris Steipe
  affiliation: 
  - http://orcid.org/0000-0002-1134-6758
  - University of Toronto
  email: boris.steipe@utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  Details on the preparation of TBD data for the BCB420.2019.ESA package.
output:
#  BiocStyle::html_document:
#    toc: true
#    toc_float: true
#    fig_caption: yes
  BiocStyle::pdf_document:
    toc: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Data Details - TBD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
bibliography: BCB420.2019.ESA-bibliography.bib
---
<!-- bibliography that appears in the references, 
but not as a citation in the body -->
nocite: @steipe-rptPlus

<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# load the package itself for knitr:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)

```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
If any of this information is ambiguous, inaccurate, outdated, or incomplete, please check the [most recent version](https://github.com/hyginn/BCB420.2019.ESA) of the package on GitHub and if the problem has not already been addressed, please 
[file an issue](https://github.com/hyginn/BCB420.2019.ESA/issues)!
</div>

&nbsp;
<!-- ======================================================================= -->

# Heading

Text





<!-- ======================================================================= -->

# Session Info

This release of the `BCB420.2019.ESA` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

<!-- the bibliography appears at the end of the page -->
# References


<!-- End -->
