---
title: "Data Details - STRINGactions"
author:
- name: Gabriela Morgenshtern
  affiliation: 
  - http://orcid.org/0000-0003-4762-8797
  - University of Toronto
  email: g.morgenshtern@mail.utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  Details on the preparation of STRINGactions protein interaction data for the BCB420.2019.ESA package.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Data Details - STRINGactions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: BCB420.2019.ESA-bibliography.bib
---

<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# load BCB420-2019.ESA itself for knitr:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)

```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
If any of this information is ambiguous, inaccurate, outdated, or incomplete, please check the [most recent version](https://github.com/hyginn/BCB420.2019.ESA) of the package on GitHub and if the problem has not already been addressed, please 
[file an issue](https://github.com/hyginn/BCB420.2019.ESA/issues)!
</div>

&nbsp;

# About this vignette:

This vignette describes the workflow that was used to prepare the STRINGactions dataset for the package. Source data is interaction types for protein links from **[STRING](https://string-db.org/) [@pmid30476243].


# STRINGactions Data

STRING is a collection of curated protein-protein interaction data. STRING protein action data is [licensed under the CC license](https://string-db.org/cgi/access.pl?footer_active_subpage=licensing). This document describes work with [STRING v11.0 protein actions data for homo sapiens (2018-11-22)](https://string-db.org/cgi/download.pl?sessionId=GiksVfWc6aQ2&species_text=Homo+sapiens) [@pmid30476243].

&nbsp;

## Data semantics

STRING interaction data is available on the STRING consortium website. 

** STRING data comes in only one format, and genes are identified by Ensemble protein IDs (ENSP). We must download this data and process it through the script below to map the ENSPs to HGNCs present a consistent identifier for our package tools to work with. 

We are interested in all the columns contained in the file `9606.protein.actions.v11.0.txt.gz`:

1. Protein 1: The ENSP identifier of the protein action pair
2. Protein 2: The ENSP identifier of the protein action pair
3. Mode: the name of the interaction, controlled vocabulary consisting of the set: {"binding",    "catalysis", "reaction", "activation", "inhibition", "ptmod", "expression"}
4. Is_directional: boolean column, answering "does one protein act on the other in this action pair?" 
5. a_is_acting: boolean column, answering: "is Protein 1 acting on Protein 2"? If false and if it is a  directional action, the opposite is true of the interaction
6. Combined_score: the [total STRING score](https://string-db.org/cgi/info.pl?footer_active_subpage=scores) computed for the protein action pair (which is in itself a STRING edge).

&nbsp;

## Data download and processing

** SEE STRING PACKAGE FOR WORDING
1. Navigate to the [download directory of the **STRING** database](https://string-db.org/cgi/download.pl).
2. Search for "human" or "homo sapiens" in the "Choose an organism" search bar
3. Download the following data file: 
* `9606.protein.actions.v11.0.txt.gz` (14.4 Mb);
3. Uncompress the file and place it into a sister directory of your working directory which is called `data`. (It should be reachable with `file.path("..", "data")`). **Warning:**  `../data/9606.protein.actions.v11.0.txt` is 211.3 Mb!

&nbsp;

### Preparations: packages, functions, files

To begin processing, we need to make sure the required packages are installed:

**`readr`** provides functions to read data which are particularly suitable for
large datasets. They are much faster than the built-in read.csv() etc. But caution: these functions return "tibbles", not data frames. ([Know the difference](https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html).)
```R
if (! requireNamespace("readr")) {
  install.packages("readr")
}
```

&nbsp;


## Selecting protein-protein interactions

```R
##### Map the protein action dataset mappings ####
tmp <- readr::read_tsv(file.path("./data", "9606.protein.actions.v11.0.txt"),
                         skip = 1,
                         col_names = c("protein1", "protein2", "mode", "action",
                                        "is_directional", "a_is_acting", "combined_score")) # 11,759,454 rows
# Keep "high confidence" interactions, and
# remove "action" col since that information is duplicated in "mode" for our purposes
tmp <- tmp[,c("protein1", "protein2", "mode",
              "is_directional", "a_is_acting", "combined_score")]
tmp <- tmp[tmp$combined_score >= 800, ]

# remove "9606." prefix
tmp$protein1 <- gsub("^9606\\.", "", tmp$protein1)
tmp$protein2 <- gsub("^9606\\.", "", tmp$protein2)

# Map ENSP to HGNC symbols: use Dr. Steipe's mapping tool:
load(file = file.path(".", "data", "ensp2sym.RData"))
tmp$protein1 <- ensp2sym[tmp$protein1]
tmp$protein2 <- ensp2sym[tmp$protein2]

# Validate initial mapping
any(grepl("ENSP", tmp$protein1))  # Nope
any(grepl("ENSP", tmp$protein2))  # None left here either

# Clean duplicate edges (from Dr. Steipe)
sPaste <- function(x, collapse = ":") {
  return(paste(sort(x), collapse = collapse))
}
tmp$key <- apply(tmp[ , c("protein1", "protein2", "mode")], 1, sPaste) # takes a min
length(tmp$key) # 2031426
length(unique(tmp$key)) # 548932
tmp <- tmp[( ! duplicated(tmp$key)),
           c("protein1", "protein2", "mode",
             "is_directional", "a_is_acting", "combined_score") ]

# Remove NA nodes
sum(is.na(tmp$protein1)) # NUM
sum(is.na(tmp$protein2)) # NUM
STRINGactions <- tmp[( ! is.na(tmp$protein1)) & ( ! is.na(tmp$protein2)), ] # 545423

# Save the file
saveRDS(STRINGactions, file = file.path(".", "data", "STRINGactions.RDS"))  # 1.6 Mb

# The dataset was uploaded to the assets server and is available with:
STRINGactions <- fetchData("STRINGactions")

```

&nbsp;

# References

&nbsp;

Steipe, Boris (2019). BCB420.2019.STRING (STRING data annotatation of human genes). [Github repository](https://github.com/hyginn/BCB420.2019.STRING)

- STRING v11 paper 
** FORMAT
Oughtred, R., Stark, C., Breitkreutz, B. J., Rust, J., Boucher, L., Chang, C., Kolas, N., O'Donnell, L., Leung, G., McAdam, R., Zhang, F., Dolma, S., Willems, A., Coulombe-Huntington, J., Chatr-Aryamontri, A., Dolinski, K., & Tyers, M. (2019). The BioGRID interaction database: 2019 update. [Nucleic acids research, D1, D529-D541](https://academic.oup.com/nar/article/47/D1/D529/5204333).

&nbsp;

# Acknowledgements

Thanks to Simon Kågedal's very useful [PubMed to APA reference tool](http://helgo.net/simon/pubmed/).

&nbsp;



# Session Info

This release of the `BCB420.2019.ESA` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

<!-- the bibliography appears at the end of the page -->
# References


<!-- End -->
