---
title: "Data Details - BioGRID"
author:
- name: Boris Steipe
  affiliation: 
  - http://orcid.org/0000-0002-1134-6758
  - University of Toronto
  email: boris.steipe@utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  Details on the preparation of BioGRID interaction data for the BCB420.2019.ESA package.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Data Details - BioGRID}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: BCB420.2019.ESA-bibliography.bib
---

<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# load BCB420-2019.ESA itself for knitr:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)

```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
If any of this information is ambiguous, inaccurate, outdated, or incomplete, please check the [most recent version](https://github.com/hyginn/BCB420.2019.ESA) of the package on GitHub and if the problem has not already been addressed, please 
[file an issue](https://github.com/hyginn/BCB420.2019.ESA/issues)!
</div>

&nbsp;

# About this vignette:

This vignette [@steipe-rptPlus] describes the workflow that was used to prepare the BioGRID dataset for the package. Source data is  protein-protein interaction data from [BioGRID ](https:/thebiogrid.org/) [@pmid30476227].


# BioGRID Data

The BioGRID is a collection of curated protein-protein interaction data. BioGRID data is [licensed under the MIT license](https://downloads.thebiogrid.org/Download/LICENSE.txt). This document describes work with [BioGRID 3.5.170 (2019-02-25)](https://downloads.thebiogrid.org/BioGRID/Release-Archive/BIOGRID-3.5.170/) [@pmid30476227].

&nbsp;

## Data semantics

BioGRID interaction data is available in various formats common to protein-protein interaction databases. For our purposes of working with HGNC symbols, the BioGRID TAB 2.0 file format appears useful. Details are described [here](https://wiki.thebiogrid.org/doku.php/biogrid_tab_version_2.0).

The file `BIOGRID-ALL-3.5.170.tab2.zip` contains the following columns of interest to us:

8. Official symbol for Interactor A. A common gene name/official symbol for interactor A. Will be a “-” if no name is available.
9. Official symbol for Interactor B. Same structure as column 8.
13. Experimental System Type. This will be either "physical" or "genetic" as a classification of the Experimental System Name.
16. Organism ID for Interactor A. This is the NCBI Taxonomy ID for Interactor A.
17. Organism ID for Interactor B. Same structure as 16.

&nbsp;

## Data download and processing

1. Navigate to the [download directory of the **BioGRID** database](https://downloads.thebiogrid.org/BioGRID/Release-Archive/BIOGRID-3.5.170/).
2. Download the following data file: 
* `BIOGRID-ALL-3.5.170.tab2.zip` (70.7 Mb);
3. Uncompress the file and place it into a sister directory of your working directory which is called `data`. (It should be reachable with `file.path("..", "data", "BioGRID")`). **Warning:**  `../data/BioGRID/BIOGRID-ALL-3.5.170.tab2.txt` is 618.9 Mb!

&nbsp;

### Preparations: packages, functions, files

To begin processing, we need to make sure the required packages are installed:

**`readr`** provides functions to read data which are particularly suitable for
large datasets. They are much faster than the built-in read.csv() etc. But caution: these functions return "tibbles", not data frames. ([Know the difference](https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html).)
```R
if (! requireNamespace("readr")) {
  install.packages("readr")
}
```

&nbsp;


## Selecting protein-protein interactions

```R
FN <- file.path("..", "data", "BioGRID", "BIOGRID-ALL-3.5.170.tab2.txt")
BioGRID <- as.data.frame(readr::read_tsv(FN,
                                         skip = 1,
                                         col_names = c("A",
                                                       "B",
                                                       "type",
                                                       "taxID_A",
                                                       "taxID_B"),
                                         col_types = "-------cc---c--ii-------"))

nrow(BioGRID)  # 1647089

# Remove all interactions for which taxID_A and taxID_B are not both 9606
BioGRID <- BioGRID[BioGRID$taxID_A == 9606 & BioGRID$taxID_B == 9606, 
                   c("A", "B", "type")]
nrow(BioGRID)  # 442753

# Remove all interactions for which A and B are not both in HGNC sym
myURL <- paste0("https://github.com/hyginn/",
                "BCB420-2019-resources/blob/master/HGNC.RData?raw=true")
load(url(myURL))  # loads HGNC data frame
BioGRID <- BioGRID[BioGRID$A %in% HGNC$sym & BioGRID$A %in% HGNC$sym, 
                   c("A", "B", "type")]
nrow(BioGRID)  # 433505

# How many genetic interactions?
sum(BioGRID$type == "genetic")  # 4702

# Save dataset:
saveRDS(BioGRID, file = file.path("..", "data", "BioGRID", "BioGRID.3.5.170.rds"))  # 2.8 Mb

# The dataset was uploaded to the assets server and is available with:
BioGRID <- fetchData("BioGRID")

```

&nbsp;

# References

&nbsp;

Oughtred, R., Stark, C., Breitkreutz, B. J., Rust, J., Boucher, L., Chang, C., Kolas, N., O'Donnell, L., Leung, G., McAdam, R., Zhang, F., Dolma, S., Willems, A., Coulombe-Huntington, J., Chatr-Aryamontri, A., Dolinski, K., & Tyers, M. (2019). The BioGRID interaction database: 2019 update. [Nucleic acids research, D1, D529-D541](https://academic.oup.com/nar/article/47/D1/D529/5204333).

&nbsp;

# Acknowledgements

Thanks to Simon Kågedal's very useful [PubMed to APA reference tool](http://helgo.net/simon/pubmed/).

&nbsp;



# Session Info

This release of the `BCB420.2019.ESA` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

<!-- the bibliography appears at the end of the page -->
# References


<!-- End -->
