---
title: "The predict_subsystem Vignette"
author:
- name: Yin Yin 
  affiliation: 
  - http://orcid.org/0000-0002-8921-6945
  - University of Toronto
  email: yin.yin@mail.utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  The Predict_subsystem Vignette is used for function findSub that is
  a tool to find new possible subsystem based on hu.map data in selectd system.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Predict_subsystem Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: BCB420.2019.ESA-bibliography.bib
---

<!--- this inline style supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# knitr knits in a new session with an empty global workspace after setting its
# working directory to ./vignettes. To make your package functions available in
# the vignette, you have to load the library. The following two lines should
# accomplish this without manual intervention:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)
```

&nbsp;

<div style="font-size:90%;border:solid 1px #337ab7;background-color:#FAFBFC;margin-right:0px;">
There are many links and references in this document. If you find anything here
ambiguous, inaccurate, outdated, incomplete, or broken, please
[file an issue](https://github.com/hyginn/rpt/issues)!
</div>

&nbsp;

# About this vignette
This vignette [@steipe-rptPlus] demonstrates the use of `findSub()` function in `BCB420.2019.ESA` package.

The function predicted new subsystem in asked system based on hu.map. The input are the system code, the database of systems and thershold for selection subsystem based on possibility score. The output is table with information of new subsystem.

The new substem is selected from overlapping enrichment and possibility score.

&nbsp;

## hu.map Data
This function use data edited from [hu.map](http://proteincomplexes.org/)[@pmid28596423] and imported by reading RMS files in data directory.

The related datas are map_complex, map_enrichment, map_prob_complex. The map_complex include all complexes and its components. The map_enrichment include enrichments for each complexes. The map_prob_complex include possibility score for expressing two proteins at the same time and involved complex code is also included for each pair.

&nbsp;


Example result table for a system:
```{r, echo=FALSE, results='asis'}
result_table <- matrix("",0,3)
colnames(result_table) <- c("System_Name", "New_Subsystem_Name", "Components")
result_table <- rbind(result_table, c("FOXDC", " Factor: FOXD2; motif: NRNWAATATTTAYN; match class: 1 ", " RPS8 RPS15A"))
knitr::kable(result_table,
             caption = "Example result Table",
             align = "c")
```


&nbsp;


# Example
## Run the function
1. check the exist system code for us that can be analyzed and choose a system as input:
```{r, eval=TRUE, include=TRUE}
names(SyDBgetRootSysIDs(fetchData("SysDB")))
```

2. Fetch the data
```{r, eval=TRUE, include=TRUE, results = "hide"}
mySDB <- fetchData("SysDB")
```

3. Run findsub by using the system we choose. In this example, I use "PHALY" and the threshold is set to be 0.5
```{r, eval=TRUE, include=TRUE, results = "hide"}
findSub("PHALY", 0.5, mySDB)
```
&nbsp;

## Results
The header of the output data frame:
```{r}
#System_Name New_Subsystem_Name                                       Components    
#[1,] "PHALY"     " Factor: FOXD2; motif: NRNWAATATTTAYN; match class: 1 " " RPS8 RPS15A"
```

&nbsp;

## Interpretation of results and future direction
The return value for the function is a table that contains system name, new subsystem name, and all components in this new subsystem.

However, in the data of this package, there is no system that can be predicted to have a new subsystem.

The new subsystem could be used to explore more about the system. It will give a new direction to explore the relationships between components in this system. Also, the function of component can be predicted also.


&nbsp;

# Session Info

This release of the `BCB420.2019.ESA` package was produced in the following context of supporting packages:

```{r, echo=TRUE}
sessionInfo()
```

<!-- the bibliography appears at the end of the page -->
# References

<!-- End -->
